Identification of protoneutron star g-modes in gravitational-wave data

### 2020-06-22 methods.tex
work on section methods

\section{Methods description}
In this section, we outline a strategy for estimating the time evolution of the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ of the mass of the proto-neutron star (PNS) and its squared radius \textcolor{red}{(units, solar Mass and km?)} based on
the gravitational wave observations.
An integral part of this strategy is the universal relationship between the characteristic frequency and the ratio of mass and radius as demonstrated by  \cite{Torres:2019}
on 25 1D simulations with the
{\sc AENUS-ALCAR }code \cite{} and the {\sc CoCoNuT} \cite{} code. Here we are using the data from \textcolor{red}{only {\sc AENUS-ALCAR} or both?, maybe colour-code the two groups of data points } to fit a cubic polynomial regression with
heteroscedastic errors
\begin{equation}
r_i=\beta_1 f_i + \beta_2 f_i^2 +\beta_3 f_i^3 + \epsilon_i
\end{equation}
where $\epsilon_i$ are assumed to be independent zero-mean Gaussian errors with  variances $\sigma_i^2$ that increase with $f_i$. The model for frequency-dependent variances is
\begin{equation}
\log \sigma_i=\alpha_0+ \alpha_1 f_i + \alpha_2 f_i^2 + \delta_i
\end{equation}
with independent and identically zero-mean Gaussian errors $\delta_i$. The R-package LMVAR
\cite{lmvar:2019} that implements a maximum likelihood approach was used to fit the model. The
best fitting model amongst polynomials of degree 1, 2, and 3  was chosen according to the  AIC, i.e. \textcolor{red}{insert correct estimates}
\begin{equation}\label{eq:universal}
r_i=f_i + f_i^3 + \epsilon_i
\end{equation}
 The best-fitting model achieves a coefficient of determination of $R^2=0.9812$.
Parameter estimates and their standard errors are given in \textcolor{red}{ add Table xxx}.
The data and fit of the model including 95\% confidence bands are displayed in Figure \ref{Fig:LMVAR}.
\begin{figure}
 \centering
 	\includegraphics[width=0.5\textwidth,height=0.3\textheight]{model.pdf}
 \caption{Data from 25 1D simulations {\sc AENUS-ALCAR } and {\sc CoCoNuT}  code, solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands.} \label{Fig:LMVAR}
\end{figure}

Here we analyse the gravitational wave signal {\tt s20-gw-10kpc} \textcolor{red}{insert more detailed description}, originally sampled at 10 kHz but resampled to the LIGO sampling rate of 16384 Hz.
A spectrogram of this signal is shown in Figure \textcolor{red}{insert spectrogram of signal} based on autoregressive estimates of the local spectra for successive time intervals of 
\textcolor{red}{length 200} with a \textcolor{red}{ 90\%} overlap.The dominant emission mode correspondes to the $\mbox{ }^2 g_1$-mode \cite{} and we have developed a time-frequency method to track the ridge $m(t)$ in the spectrogram, taking into account that it is monotonically increasing. Starting from either the left- or right-most column of the time-frequency  matrix we identify and trace the sequence of amplitude peaks within a certain frequency band given the monotonicity constraint. \textcolor{red}{Are more details regarding the ridge tracking required here?}

>>>>

\section{Methods description}
In this section, we outline a strategy for estimating the time evolution of the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ of the mass of the proto-neutron star (PNS) and its squared radius (in units of solar mass and km) from the gravitational wave observations.
An integral part of this strategy is the universal relationship between the characteristic frequency of the PNS oscillation $f$, $g$ and $p$ modes and a combination of the mass, the radius of the PNS, the shock radius and the total mass inside the shock as demonstrated by \cite{Torres:2019}. In this paper we focus on the $^2g_2$ mode.

Using 25 1D simulations obtained with the {\sc AENUS-ALCAR }code \cite{} and the {\sc CoCoNuT} \cite{} code, we parametrize the ratio with a cubic polynomial regression with heteroscedastic errors
%Here we are using the data from \textcolor{red}{only {\sc AENUS-ALCAR} or both?, maybe colour-code the two groups of data points } to fit a cubic polyn
\begin{equation}
r_i=\beta_1 f_i + \beta_2 f_i^2 +\beta_3 f_i^3 + \epsilon_i
\end{equation}
where $\epsilon_i$ are assumed to be independent zero-mean Gaussian errors with  variances $\sigma_i^2$ that increase with frequency $f_i$. The model for frequency-dependent variances is
\begin{equation}
\log \sigma_i=\alpha_0+ \alpha_1 f_i + \alpha_2 f_i^2 + \delta_i
\end{equation}
with independent and identically zero-mean Gaussian errors $\delta_i$. The R-package LMVAR
\cite{lmvar:2019} that implements a maximum likelihood approach was used to fit the model. The best
fitting model amongst polynomials of degree 1, 2, and 3  was chosen according to the Aikaike information
criterion with coefficients given in Table \ref{tab:model},

\begin{equation}\label{eq:universal}
r_i = \beta_1 f_i + \beta_3 f_i^3 + \epsilon_i
\end{equation}

The best-fitting model achieves a coefficient of determination of $R^2=0.9812$.
The data and fit of the model including 95\% confidence bands are displayed in Figure \ref{Fig:LMVAR}.

\begin{table}[h]
  \begin{tabular}{lll}
    \hline
    Coefficient & Estimate & standard error \\
    \hline
    $\beta_1$   & $6.09 \times 10^{-7}$ & $1.75 \times 10^{-8}$ \\
    $\beta_3$   & $6.24 \times 10^{-13}$ & $8.79 \times 10^{-15}$ \\
    \hline
  \end{tabular}
\label{tab:model}
\caption{Estimate and standard error of the coefficients of the best fit model describing the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ as function of the frequency of the $^2g_2$ mode.}
\end{table}

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth,height=0.3\textheight]{model.pdf}
 \caption{Data from 25 1D simulations {\sc AENUS-ALCAR } and {\sc CoCoNuT}  code, solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands.} \label{Fig:LMVAR}
\end{figure}

We consider the gravitational wave signal {\tt s20-gw-10kpc} described in Section \ref{sec:simulation}, originally sampled at 20 kHz but resampled to the LIGO sampling rate of 16384 Hz.
A spectrogram of this signal is shown in Figure \ref{fig:spectrogram} based on autoregressive estimates of the local spectra for successive time intervals of 
\textcolor{red}{length 200} with a \textcolor{red}{ 90\%} overlap.The dominant emission mode correspondes to the $\mbox{ }^2 g_1$-mode \cite{} and we have developed a time-frequency method to track the ridge $m(t)$ in the spectrogram, taking into account that it is monotonically increasing. Starting from either the left- or right-most column of the time-frequency  matrix we identify and trace the sequence of amplitude peaks within a certain frequency band given the monotonicity constraint. \textcolor{red}{Are more details regarding the ridge tracking required here?}

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth,height=0.3\textheight]{plots/spectrogram.pdf}
 \caption{Spectrogram} \label{fig:spec}
\end{figure}
